Further Modeling of $q_{95}$ Windows for the Suppression of Edge Localized Modes by Resonant Magnetic Perturbations in the DIII-D Tokamak

### 2020-09-10 Papers/Paper2/Paper.tex
Minor corrections to Paper2

Equation~(\ref{e5}) is usually coupled with the so-called ``no-slip constraint'':\,\cite{rfa}
\begin{equation}\label{e6}
\frac{d\varphi_s}{dt} = \omega_s.
\end{equation}
The reasoning behind the imposition of this constraint is that the plasma is trapped inside the
magnetic separatrix of the  island chain that forms in the vicinity of the rational surface,
which forces the chain to co-rotate with the local plasma flow. However, in reality, the plasma
is able to diffuse resistively across the separatrix to some extent. Hence, by analogy with Eqs.~(\ref{e3}), (\ref{e4}), and (\ref{e5}), and 
in accordance with Refs.~\onlinecite{slip1} and \onlinecite{slip2}, in this paper we shall modify Eq.~(\ref{e6}) such that
\begin{equation}\label{e7}
\frac{d\varphi_s}{dt} = \omega_s - \frac{E_{sv}}{\tau_R}\,\frac{r_s}{({\cal I}/2)\,W_s}\,\frac{|{\mit\Psi}_v|}{|{\mit\Psi}_s|}\,\sin(\varphi_s-\varphi_v).
\end{equation}
The final term on the right-hand side of the previous equation is the ``nonlinear slip-frequency''. In general,
we would expect this frequency to be relatively small (because it is usually the case that $|\omega_s|\,\tau_R\,(W_s/r_s)\gg 1$
for fully-developed magnetic islands in high-temperature tokamak plasmas\,\cite{rf1,slip2}). Nevertheless, the
nonlinear slip-frequency may be non-negligible for developing island chains, which is why we are including it in our analysis.
(Note that the nonlinear slip-frequency does not appear in rigorous island calculations, such as
that described in Ref.~\onlinecite{rfisland}, and references therein, because such calculations depend crucially on
complicated hierarchical ordering assumptions, according to which the nonlinear slip-frequency is negligible.)

@@

Equation~(D6) of Ref.~\onlinecite{rftor} is modified such that
\begin{align}\label{e19}
\left(\hat{W}_k +\hat{\delta}_k\right){\cal S}_k\left(\frac{d\varphi_k}{d\hat{t}}-\hat{\varpi}_k\right)&=
- \sum_{k'=1,K}\hat{E}_{kk'}\,\hat{\mit\Psi}_{k'}\,\sin(\varphi_k-\varphi_{k'}-\xi_{kk'})-\hat{E}_{kk}\,\hat{\chi}_k\,\sin(\varphi_k-\zeta_k),
\end{align}
where
\begin{equation}
\hat{\varpi}_k= \hat{\varpi}_{k\,0}+m_k\,{\mit\Delta}\hat{\mit\Omega}_\theta(\hat{r}_k,\hat{t})-n\,{\mit\Delta}\hat{\mit\Omega}_\phi(\hat{r}_k,\hat{t}),
\end{equation}
and $\hat{\varpi}_{k\,0}= \varpi_{k\,0}\,\tau_A$, $\hat{\mit\Omega}_\theta= {\mit\Delta}{\mit\Omega}_\theta\,\tau_A$, and $\hat{\mit\Omega}_\phi= {\mit\Delta}{\mit\Omega}_\phi\,\tau_A$. Here, $\varpi_{k\,0}$
is the so-called ``natural frequency'' at the $k$th rational surface; this quantity is defined as the helical phase velocity of a naturally unstable island chain, resonant at the surface, in the
absence of an RMP (or any other island chains). On the other hand,  ${\mit\Delta}{\mit\Omega}_\theta$ and ${\mit\Delta}{\mit\Omega}_\phi$ are the changes in the plasma
poloidal and toroidal velocity profiles, respectively, induced by the electromagnetic torques that develop in the vicinities of the various rational surfaces in the plasma in response to the RMP.
Equation~(\ref{e19}) represents
a major improvement to Eq.~(D6) of Ref.~\onlinecite{rftor} that takes into  account the finite slip-frequencies of the magnetic island chains induced at
the various rational surfaces in the plasma. In the absence of a finite slip-frequency, the right-hand side of Eq.~(\ref{e19}) would be zero.

@@

\subsection{Relaxation of No-Slip Constraint}\label{s4a}
As is discussed in Ref.~\onlinecite{rftor1}, the best agreement between the predictions of the {\tt EPEC} code and experimental observations is
obtained by assuming that the natural frequencies at the various rational surfaces in the plasma are determined by the
local equilibrium ${\bf E}\times {\bf B}$ frequencies. In other words, 
\begin{equation}\label{e24}
\varpi_{k\,0} =-n\,\omega_E({\mit\Psi}_{p\,k}),
\end{equation}
where 
\begin{equation}
\omega_E({\mit\Psi}_p) = -\frac{d{\mit\Phi}}{d{\mit\Psi}_p}.
\end{equation}
Here, ${\mit\Phi}({\mit\Psi}_p)$ is the equilibrium scalar electric potential, ${\mit\Psi}_p$ the equilibrium poloidal magnetic flux (divided by $2\pi$), 
and ${\mit\Psi}_{p\,k}$  the value of ${\mit\Psi}_p$ at the $k$th rational surface.  Equation~(\ref{e24}) holds at
every rational surface in all of the simulations discussed in Sect.~\ref{s4a}.

@@

A comparison of Figs.~\ref{fig2} and \ref{fig4}, as well as an examination of Fig.~\ref{fig6}, reveals that the relaxation of the
no-slip constraint facilitates the locking of magnetic island chains to the RMP to a relatively modest degree. (Here, by locking, we mean a reduction in the
RMP-modified natural frequency to a value close to zero.) 
This is not surprising, since
relaxation of the no-slip constraint clearly weakens the coupling between a given magnetic island chain and the local plasma flow (which is ultimately responsible
for the shielding of driven magnetic reconnection). 
It can be seen from Fig.~\ref{fig6} that if the no-slip constraint is imposed then the RMP-modified natural frequencies of locked magnetic
island chains are exactly zero. On the other hand, if the no-slip constraint is relaxed then locked magnetic island chains possess
 small but finite RMP-modified natural frequencies, because the local plasma flow is able to slip through the  chains to some
 extent. However, the degree of slippage for a fully developed locked magnetic island chain is comparatively small (i.e., no more than
 $0.5\,{\rm krad/s}$), which helps explain
 why the no-slip constraint can be used to accurately predict the dynamics of wide magnetic island chains in tokamaks and reversed field pinches.\cite{rfp}

@@

\subsection{Improved Interpolation between Linear and Nonlinear Resonant Response Regimes}\label{s4b}
According to linear tearing mode theory, in the absence of the RMP, the natural frequency of the tearing mode resonant at the $k$th rational surface is given by\,\cite{ara,lin1,lin2,lin3}
\begin{equation}\label{elin}
\varpi_{k\,0} = \varpi_{{\rm ln}\,k} \equiv - n\,(\omega_E+\omega_{\ast\,e})_{{\mit\Psi}_{p\,k}},
\end{equation}
where
\begin{align}
\omega_{\ast\,e}({\mit\Psi}_p) &=\frac{T_e}{e}\,\frac{d\ln p_e}{d{\mit\Psi}_p}.
\end{align}
 Here,  $e$ is the magnitude of the electron charge,  $p_e({\mit\Psi}_p)$  the equilibrium
electron pressure, and $T_e({\mit\Psi}_p)$ the equilibrium
electron temperature.

>>>>

Equation~(\ref{e5}) is usually coupled with the so-called ``no-slip constraint'':\,\cite{rfa}
\begin{equation}\label{e6}
\frac{d\varphi_s}{dt} = \omega_s.
\end{equation}
The reasoning behind the imposition of this constraint is that  plasma is trapped inside the
magnetic separatrix of the  island chain that forms in the vicinity of the rational surface,
which forces the chain to co-rotate with the local plasma flow. However, in reality, the plasma
is able to diffuse resistively across the separatrix to some extent. Hence, by analogy with Eqs.~(\ref{e3}), (\ref{e4}), and (\ref{e5}), and 
in accordance with Refs.~\onlinecite{slip1} and \onlinecite{slip2}, in this paper we shall modify Eq.~(\ref{e6}) such that
\begin{equation}\label{e7}
\frac{d\varphi_s}{dt} = \omega_s - \frac{E_{sv}}{\tau_R}\,\frac{r_s}{({\cal I}/2)\,W_s}\,\frac{|{\mit\Psi}_v|}{|{\mit\Psi}_s|}\,\sin(\varphi_s-\varphi_v).
\end{equation}
The final term on the right-hand side of the previous equation is the ``nonlinear slip-frequency''. In general,
we would expect this frequency to be relatively small (because it is usually the case that $|\omega_s|\,\tau_R\,(W_s/r_s)\gg 1$
for fully-developed magnetic islands in high-temperature tokamak plasmas\,\cite{rf1,slip2}). Nevertheless, the
nonlinear slip-frequency may be non-negligible for developing island chains, which is why we are including it in our analysis.
(Note that the nonlinear slip-frequency does not appear in rigorous island calculations, such as
that described in Ref.~\onlinecite{rfisland}, and references therein, because such calculations depend crucially on
complicated hierarchical ordering assumptions, according to which the nonlinear slip-frequency is negligible.)

@@

Equation~(D6) of Ref.~\onlinecite{rftor} is modified such that
\begin{align}\label{e19}
\left(\hat{W}_k +\hat{\delta}_k\right){\cal S}_k\left(\frac{d\varphi_k}{d\hat{t}}-\hat{\varpi}_k\right)&=
- \sum_{k'=1,K}\hat{E}_{kk'}\,\hat{\mit\Psi}_{k'}\,\sin(\varphi_k-\varphi_{k'}-\xi_{kk'})-\hat{E}_{kk}\,\hat{\chi}_k\,\sin(\varphi_k-\zeta_k),
\end{align}
where
\begin{equation}
\hat{\varpi}_k= \hat{\varpi}_{k\,0}+m_k\,{\mit\Delta}\hat{\mit\Omega}_\theta(\hat{r}_k,\hat{t})-n\,{\mit\Delta}\hat{\mit\Omega}_\phi(\hat{r}_k,\hat{t}),
\end{equation}
and $\hat{\varpi}_{k\,0}= \varpi_{k\,0}\,\tau_A$, $\hat{\mit\Omega}_\theta= {\mit\Delta}{\mit\Omega}_\theta\,\tau_A$, and $\hat{\mit\Omega}_\phi= {\mit\Delta}{\mit\Omega}_\phi\,\tau_A$. Here, $\varpi_{k\,0}$
is the so-called ``natural frequency''  (in the absence of the RMP) at the $k$th rational surface; this quantity is defined as the helical phase velocity of a naturally unstable island chain, resonant at the surface, in the
absence of an RMP (or any other island chains). On the other hand,  ${\mit\Delta}{\mit\Omega}_\theta$ and ${\mit\Delta}{\mit\Omega}_\phi$ are the changes in the plasma
poloidal and toroidal velocity profiles, respectively, induced by the electromagnetic torques that develop in the vicinities of the various rational surfaces in the plasma in response to the RMP.
Equation~(\ref{e19}) represents
a major improvement to Eq.~(D6) of Ref.~\onlinecite{rftor} that takes into  account the finite slip-frequencies of the magnetic island chains induced at
the various rational surfaces in the plasma. In the absence of a finite slip-frequency, the right-hand side of Eq.~(\ref{e19}) would be zero.

@@

\subsection{Relaxation of No-Slip Constraint}\label{s4a}
As is discussed in Ref.~\onlinecite{rftor1}, the best agreement between the predictions of the {\tt EPEC} code and experimental observations is
obtained by assuming that the unperturbed natural frequencies at the various rational surfaces in the plasma are determined by the
local equilibrium ${\bf E}\times {\bf B}$ frequencies. In other words, 
\begin{equation}\label{e24}
\varpi_{k\,0} =-n\,\omega_E({\mit\Psi}_{p\,k}),
\end{equation}
where 
\begin{equation}
\omega_E({\mit\Psi}_p) = -\frac{d{\mit\Phi}}{d{\mit\Psi}_p}.
\end{equation}
Here, ${\mit\Phi}({\mit\Psi}_p)$ is the equilibrium scalar electric potential (in the absence of the RMP), ${\mit\Psi}_p$ the equilibrium poloidal magnetic flux (divided by $2\pi$), 
and ${\mit\Psi}_{p\,k}$  the value of ${\mit\Psi}_p$ at the $k$th rational surface.  Equation~(\ref{e24}) holds at
every rational surface in all of the simulations discussed in Sect.~\ref{s4a}.

@@

A comparison of Figs.~\ref{fig2} and \ref{fig4}, as well as an examination of Fig.~\ref{fig6}, reveals that the relaxation of the
no-slip constraint facilitates the locking of magnetic island chains to the RMP to a relatively modest degree. (Here, by locking, we mean a reduction in the
RMP-modified natural frequency to a value close to zero.) 
This is not surprising, since
relaxation of the no-slip constraint clearly weakens the coupling between a given magnetic island chain and the local plasma flow (which is ultimately responsible
for the shielding of driven magnetic reconnection). 
It can be seen from Fig.~\ref{fig6} that if the no-slip constraint is imposed then the RMP-modified natural frequencies of locked magnetic
island chains are exactly zero. On the other hand, if the no-slip constraint is relaxed then locked magnetic island chains possess
 small but finite RMP-modified natural frequencies, because the local plasma flow is able to slip through the  stationary chains to some
 extent. However, the degree of slippage for a fully developed locked magnetic island chain is comparatively small (i.e., no more than
 $0.5\,{\rm krad/s}$), which helps explain
 why the no-slip constraint can be used to accurately predict the dynamics of wide magnetic island chains in tokamaks and reversed field pinches.\cite{rfp}

@@

\subsection{Improved Interpolation between Linear and Nonlinear Resonant Response Regimes}\label{s4b}
According to linear tearing mode theory, in the absence of the RMP, the natural frequency of the tearing mode resonant at the $k$th rational surface is given by\,\cite{ara,lin1,lin2,lin3}
\begin{equation}\label{elin}
\varpi_{k\,0} = \varpi_{{\rm ln}\,k} \equiv - n\,(\omega_E+\omega_{\ast\,e})_{{\mit\Psi}_{p\,k}},
\end{equation}
where
\begin{align}
\omega_{\ast\,e}({\mit\Psi}_p) &=\frac{T_e}{e}\,\frac{d\ln p_e}{d{\mit\Psi}_p}.
\end{align}
 Here,  $e$ is the magnitude of the electron charge,  $p_e({\mit\Psi}_p)$  the equilibrium (i.e., in the absence of the RMP)
electron pressure, and $T_e({\mit\Psi}_p)$ the equilibrium
electron temperature.
